Research log

### 2023-01-28 docs/ExploratoryResults.Rmd
simulation or pretense in battleship and hangman (RT and optimality analysis)

N_perm <- 1000;
bootstrap_error <- function(x, N_perm) {
  N <- length(x)
  medians = c();
  for (i in 1:N_perm) {
    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median())
  };
  return(sd(medians))
}

@@

```{r suppression-or-simulation}

@@

In the following analysis, we focus on the first two labels and carve out two sets of players: `r E2.suppressors%>%length` players who reported using suppression but not simulation, and `r E2.simulators%>%length` who reported using simulation but no suppression.

@@

```
The mean negative optimality score of non-pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)))$statistic`). The mean negative optimality score of pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='pretend') %>%pull(p_click_rank)%>%mean()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='pretend') %>%pull(p_click_rank)%>%mean()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='pretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='pretend') %>%pull(p_click_rank)))$statistic`). To the very least, simulators did as well as suppressors in pretending.

@@

>>>>

N_perm <- 1000;
bootstrap_error <- function(x, N_perm) {
  N <- length(x)
  medians = c();
  for (i in 1:N_perm) {
    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median(na.rm=T))
  };
  return(sd(medians,na.rm=T))
}

@@

```{r battleship-suppression-or-simulation}

@@

In the following analysis, we focus on the first two labels and carve out two sets of players: `r E2.suppressors%>%length` players who reported using suppression but not simulation, and `r E2.simulators%>%length` who reported using simulation but not suppression.

@@

The mean negative optimality score of non-pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)))$statistic`). The mean negative optimality score of pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='pretend') %>%pull(p_click_rank)%>%mean()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='pretend') %>%pull(p_click_rank)%>%mean()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='pretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='pretend') %>%pull(p_click_rank)))$statistic`). To the very least, simulators did as well as suppressors in pretending.

### RT by outcome

```{r RT_by_accuracy-sup-sim}

E2.RT_by_lag3 <- E2.click_log %>%
  filter(click_number>3) %>%
  group_by(subj_id,test_part,lag3,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lag3,RT,sep='_') %>%
  mutate(lag3_diff = lag3_TRUE-lag3_FALSE);

E2.RT_by_lag2 <- E2.click_log %>%
  filter(click_number>2) %>%
  group_by(subj_id,test_part,lag2,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lag2,RT,sep='_') %>%
  mutate(lag2_diff = lag2_TRUE-lag2_FALSE);

E2.RT_by_lag1 <- E2.click_log %>%
  filter(click_number>1) %>%
  group_by(subj_id,test_part,lag1,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lag1,RT,sep='_') %>%
  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)

E2.RT_by_hit <- E2.click_log %>%
  filter(click_number>0) %>%
  group_by(subj_id,test_part,hit_bin,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(hit_bin,RT,sep='_') %>%
  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)

E2.RT_by_lead1 <- E2.click_log %>%
  filter(click_number<num_clicks) %>%
  group_by(subj_id,test_part,lead1,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lead1,RT,sep='_') %>%
  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);

E2.RT_by_lead2 <- E2.click_log %>%
  filter(click_number<num_clicks-1) %>%
  group_by(subj_id,test_part,lead2,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lead2,RT,sep='_') %>%
  mutate(lead2_diff = lead2_TRUE-lead2_FALSE)

E2.RT_by_lead3 <- E2.click_log %>%
  filter(click_number<num_clicks-2) %>%
  group_by(subj_id,test_part,lead3,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lead3,RT,sep='_') %>%
  mutate(lead3_diff = lead3_TRUE-lead3_FALSE)

E2.RT_by_hit_lags <- E2.RT_by_lag3 %>%
  merge(E2.RT_by_lag2) %>%
  merge(E2.RT_by_lag1) %>%
  merge(E2.RT_by_hit) %>%
  merge(E2.RT_by_lead1) %>%
  merge(E2.RT_by_lead2) %>%
  merge(E2.RT_by_lead3);

E2.RT_by_hit_lags_long <- E2.RT_by_hit_lags %>%
  gather("position","difference",c("lag3_diff","lag2_diff", "lag1_diff", "hit_diff", "lead1_diff", "lead2_diff", "lead3_diff"))%>%
  mutate(position = ifelse(position=="lag3_diff", 3,
                           ifelse(position=="lag2_diff",2,
                                  ifelse(position=="lag1_diff",1,
                                         ifelse(position=="hit_diff",0,
                                                ifelse(position=="lead1_diff",-1,
                                                       ifelse(position=="lead2_diff",-2,-3
                                                              ))))))) %>%
  dplyr::select(subj_id,test_part,genuine_first,position,difference);

E2.RT_by_hit_lags_long %>% 
  filter(position<2 & position > -2)%>%
  filter(subj_id %in% E2.simulators | subj_id %in% E2.suppressors) %>%
  mutate(group = ifelse(subj_id %in% E2.simulators, 
                        'simulators','suppressors')) %>%
  group_by(test_part,position, group) %>%
  summarise(median_difference=median(difference),
            sem_difference=bootstrap_error(difference,N_perm)) %>%
  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  scale_x_continuous(breaks=-1:1,labels=c('hit on next trial', 'hit on current trial', 'hit on previous trial'))+
  geom_abline(slope=0,intercept=0,linetype=3)+
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
  theme_classic() +
  facet_wrap(~group, nrow=2)

ggsave('figures/E2_RT_by_lag_sim_sup.pdf',width=7)
ggsave('figures/E2_RT_by_lag.png_sim_sup.png',width=4, height=3.5, dpi=300)

Visual inspection reveals similar temporal profiles in both groups (the higher standard error among suppressors is due to the lower number of subjects in this group). Specifically, in pretend games, cell selections that resulted in a hit were faster than those that resulted in a miss in both groups (suppressors: `r apa_print(E2.RT_by_hit_lags %>% filter(test_part=='pretend' & subj_id %in% E2.suppressors)%>%pull(hit_diff)%>%t.test())$full_result`; simulators:  `r apa_print(E2.RT_by_hit_lags %>% filter(test_part=='pretend' & subj_id %in% E2.simulators)%>%pull(hit_diff)%>%t.test())$full_result`), and cell selections that followed a hit were slower than those that followed a miss in both groups (suppressors: `r apa_print(E2.RT_by_hit_lags %>% filter(test_part=='pretend' & subj_id %in% E2.suppressors)%>%pull(lag1_diff)%>%t.test())$full_result`; simulators:  `r apa_print(E2.RT_by_hit_lags %>% filter(test_part=='pretend' & subj_id %in% E2.simulators)%>%pull(lag1_diff)%>%t.test())$full_result`).

@@

```{r hangman-suppression-or-simulation}

At the end of the experiment, players were asked whether they had a strategy for pretending or for detecting pretense. I (MM) manually labeled their responses according to three criteria: mentions of suppression (e.g., "Trying to ignore the hints"), simulation (e.g., "Trying to imagine how I would react"), and rules (e.g., "Checking the corners and center before random selection of tiles"). Some answers got more than one label and some got none. Overall, `r E4.debrief_labeled$suppression%>%sum()` mentioned using suppression, `r E4.debrief_labeled$simulation%>%sum()` mentioned using simulation, and `r E4.debrief_labeled$heuristic%>%sum()` mentioned following a heuristic.

Since there were only `r E4.suppressors%>%length` players who reported using suppression but not simulation, we focus our analysis only on the `r E4.simulators%>%length` who reported using simulation but no suppression.

```{r battleship-p_click-sim, echo = FALSE, cache=TRUE}

E4.sim.mean_P_click_rank <- E4.click_log_with_boards %>%
  rbind(E4.click_log_with_boards_random) %>%
  ungroup() %>%
  filter(as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators)) %>%
  group_by(subj_id,test_part) %>%
  summarise(p_click_rank=mean(p_click_rank));

E4.sim.mean_P_click_rank_misses_only <- E4.click_log_with_boards %>%
  rbind(E4.click_log_with_boards_random) %>%
  ungroup() %>%
  filter(as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators) & !hit_bin) %>%
  group_by(subj_id,test_part) %>%
  summarise(p_click_rank=mean(p_click_rank));

E4.sim.mean_P_click_rank %>%
    filter(test_part != 'optimal') %>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(letter in word) - rank', y='number of players', title='Simulators only')+
    theme(legend.position=c(0.10,0.85)) +
    # scale_x_reverse() +
    theme_classic();

ggsave('figures/E4_pclick_rank_sim_only.pdf')
ggsave('figures/E4_pclick_rank_sim_only.png',width=5,height=2,dpi=300)

p<- E4.sim.mean_P_click_rank_misses_only %>%
    filter(test_part != 'optimal') %>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(letter in word) - rank', y='number of players', title='Simulators only: misses')+
    theme(legend.position=c(0.10,0.85)) +
    # scale_x_reverse() +
    theme_classic();

ggsave('figures/E4_pclick_rank_sim_only_misses_only.pdf',p)
ggsave('figures/E4_pclick_rank_sim_only_misses_only.png',p,width=5,height=2,dpi=300)

Among simulators, pretend games were significantly more optimal than non-pretend games (`r apa_print(E4.sim.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test())$full_result`) even when restricting the analysis to letter selections that resulted in a miss (`r apa_print(E4.sim.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test())$full_result`). Like in the full sapmle, simulators were more optimal when playing normally than when pretending (`r apa_print(E4.sim.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`), also when focusing on misses (`r apa_print(E4.sim.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).

### RT by outcome

```{r hangman-RT-by-accuracy-sim}

E4.click_log <- E4.click_log %>%
  group_by(subj_id,test_part,word) %>%
 mutate(hit_bin = hit !=0,
        lag3 = lag(hit,3)!=0,
        lag2 = lag(hit,2)!=0,
        lag1 = lag(hit,1)!=0,
        lead1 = lead(hit,1)!=0,
        lead2 = lead(hit,2)!=0,
        lead3 = lead(hit,3)!=0)

E4.RT_by_lag1 <- E4.click_log %>%
  filter(click_number>1 & click_number<num_clicks-1) %>%
  group_by(genuine_first,subj_id,test_part,lag1)%>%
  summarise(RT=median(RT, na.rm=T))  %>%
  spread(lag1,RT,sep='_') %>%
  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)

E4.RT_by_hit <- E4.click_log %>%
  filter(click_number>1 & click_number<num_clicks-1) %>%
  group_by(genuine_first,subj_id,test_part,hit_bin)%>%
  summarise(RT=median(RT, na.rm=T))  %>%
  spread(hit_bin,RT,sep='_') %>%
  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)

E4.RT_by_lead1 <- E4.click_log %>%
  filter(click_number>1 & click_number<num_clicks-1) %>%
  group_by(genuine_first,subj_id,test_part,lead1)%>%
  summarise(RT=median(RT, na.rm=T))  %>%
  spread(lead1,RT,sep='_') %>%
  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);

E4.RT_by_hit_lags <- E4.RT_by_lag1 %>%
  merge(E4.RT_by_hit) %>%
  merge(E4.RT_by_lead1)

E4.RT_by_hit_lags_long <- E4.RT_by_hit_lags %>%
  gather("position","difference",c("lag1_diff", "hit_diff", "lead1_diff"))%>%
  mutate(position = ifelse(position=="lag1_diff",1,
                                         ifelse(position=="hit_diff",0,-1))) %>%
  dplyr::select(genuine_first,subj_id,test_part,position,difference);

E4.RT_by_hit_lags_long %>%
  filter(as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators)) %>%
  group_by(test_part,position) %>%
  summarise(median_difference=median(difference, na.rm=T),
            sem_difference=bootstrap_error(difference,N_perm)) %>%
  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  scale_x_continuous(breaks=-3:3,labels=c('t-3','t-2','t-1','t','t+1','t+2','t+3'))+
  geom_abline(slope=0,intercept=0,linetype=3)+
  geom_vline(xintercept=0,linetype=3)+
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
  labs(fill=' ',x='Relative serial position', y='RT difference (ms) between hit or miss on timepoint t', title='Simulators only')+
  theme_classic()

ggsave('figures/Hangman2/RT_by_lag_acc_sim.pdf',width=7)
ggsave('figures/Hangman2/RT_by_lag_acc_sim.png',width=5, height=3, dpi=300)

E4.sim.RT_by_hit_lags_by_condition <- E4.RT_by_hit_lags_long %>%
  filter(as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators)) %>%
  spread(test_part,difference) %>%
  mutate(interaction = pretend-nonpretend)

The simulators group showed a similar temporal profile to that of the entire sample. Specifically, in pretend games, cell selections that resulted in a hit were faster than those that resulted in a miss (`r apa_print(E4.RT_by_hit_lags %>% filter(test_part=='pretend' & as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators))%>%pull(hit_diff)%>%t.test())$full_result`), and cell selections that followed a hit were slower than those that followed a miss (`r apa_print(E4.RT_by_hit_lags %>% filter(test_part=='pretend' & as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators))%>%pull(lag1_diff)%>%t.test())$full_result`).

### 2023-01-28 docs/preregisteredMethodsAndResults.Rmd
simulation or pretense in battleship and hangman (RT and optimality analysis)

---
title             : "Pretending not to know"
shorttitle        : "Pretending not to know"

@@

E4.mean_P_click_rank %>%
    filter(test_part != 'optimal') %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="black", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    # facet_wrap(~genuine_first, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

```
### Hypothesis 7b (Letter frequency)

@@

E4.mean_letter_prob_P_click_rank %>%
    filter(test_part != 'optimal') %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="black", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    # facet_wrap(~genuine_first, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

>>>>

---
title             : "Pretending not to know: pre-registered analyses"
shorttitle        : "Pretending not to know"

@@

E4.mean_P_click_rank %>%
    filter(test_part != 'optimal') %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="black", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(letter in word) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    # facet_wrap(~genuine_first, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

### Hypothesis 7b (Letter frequency)

@@

E4.mean_letter_prob_P_click_rank %>%
    filter(test_part != 'optimal') %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="black", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(letter in English text) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    # facet_wrap(~genuine_first, nrow=2)+
    # scale_x_reverse() +
    theme_classic();
